---
title: "Hypothetical Design"
---

```{r}

library(cli)
library(tidyverse)
library(data.table)
library(theme61)
library(tidyr)
library(readxl)
library(gghighlight)
library(readabs)
library(OECD)
library(jsonlite)
library(httr)
library(Synth)
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

------------------------------------------------------------------------

# Hypothetical design

## Insurance or a flat payment

The Australian social security system was based on a similar principle to systems implements in the United Kingdom and New Zealand at the same time. Following the Beveridge Report's indication for the need for **social insurance** to prevent hardship and encourage participation following World War II, Australia introduced the Social Security Act.

This act established a **flat rate** and **non-contributory** payment, with the rate paid varying due to the nature of the hardship experienced. This also took responsibility for such schemes largely away from individual States and thereby standardised treatment across Australia.

The element of unemployment benefits that reflects a flat rate, non-contributory, payment is typically called **unemployment assistance** (UA) [@oecd2024]. XXX

Although this sounds similar to a **universal basic income** (UBI) there are key distinctions:

1.  A UBI is paid to every individual, while UA is contingent on family structure and income.

2.  A UBI is not directly abated at an individuals income increases - as a result for the same tax scale effective marginal tax rates are much lower for UBI recipients.

3.  Commonly, a UBI does not involve mutual obligations or a specific requirement for job search while UA generally does.

Another element of any social support system is **unemployment insurance** (UI). The insurance element of a social support system differs from the assistance element in terms of the payment (typically linked to income prior to the event) and the funding mechanism (typically involves a contribution from the individual prior to the event which determines their eligibility).

```{r}
#| echo: false

data <- data.frame(
  Treatment = c("Payment relative to income", "Employee contributions", "Individual Means testing","Family based Means testing","Job Search obligations","Time limitations"),
  `Unemployment Assistance` = c("Flat Rate", "None", "Value 8", "Value 8", "Value 8","Limited"),
  `Univeral Basic Income` = c("Flat Rate", "None", "Value 9", "Value 8", "Value 8","None"),
  `Unemployment Insurance` = c("Relative to prio Income", "Levy", "Value 9", "Value 8", "Value 8","Between X and X")
)

# Generate the table
kable(data, caption = "Example Table Using R")
```

### Broad rationale for distinction in payments

The reason why such policies are considered as distinct, even though they form part of the same social safety net, is that the policy objectives of each part of the system differ. A UA/UBI payment reflects a **minimum income standard** related to views around poverty. A UI system reflects an **insurance** motive where individuals can pool ex-ante risk relating to a shock.

In terms of the **minimum income standard** this reflects in the necessary income that is required to meet this standard. The nature of the life event (i.e. disability, temporary illness, redundancy, or voluntary unemployment) and the structure of the family

In terms of **insurance** this reflects in XXX

In this note we are focused solely on ***job loss risk***. Such an event is deemed to be **temporary** and therefore XXX

### The income unit

The income unit reflects a set of individuals who share resources, and so may be able to consume more than sum of their private incomes would suggest. The primary sharing that is often considered is intra-household sharing, or what is termed **economies of scale in the household**.

## The Iron Triangle

The key policy framework used to evaluate the trade-offs inherent in setting benefit payment and abatement rates is the **Iron Triangle of Welfare** (cite Henry Tax Review, McClure Report, Welfare Working Group report, and Blundell Keynes Lecture 2021).

```{r}
#| echo: false

# Define the vertices of the Iron Triangle
triangle_data <- data.frame(
  x = c(0, 1, 0.5),
  y = c(0, 0, sqrt(3)/2),
  label = c("Adequate Income", "Limited Fiscal Cost", "Work Incentives")
)

ggplot(triangle_data, aes(x = x, y = y)) +
  geom_polygon(fill = "lightblue", color = "black") +
  geom_text(aes(label = label, vjust = ifelse(label == "Work Incentives", -1, 2.5)), 
            hjust = 0.5, size = 8) +
  theme_void() +
  coord_fixed() +
  xlim(-0.5, 1.5) + 
  ylim(-0.5, 1.2)    

```

In terms of objectives, the Iron Triangle is predicated on providing a minimum income standard without "discouraging self-sufficiency" - an objective that is often termed limited disincentives to work.

Targeting efficiency XXX

\[Fiscal cost\]

\[Poverty rates\]

\[Venn diagrams\]

## Rationale for varying systems

Above we noted that the distinguishing policy rationale for the unemployment assistance (UA) attribute of a support system and the unemployment insurance (UI) attribute was a **minimum income standard** as opposed to **lifetime income insurance**. In this section we will be a bit more specific with this distinction.

```{r}

#| echo: false

data <- data.frame(
  `Policy Aim` = c("Welfare-to-work", "Income Adequacy", "Fiscal Cost","XXX","XXX","XXX"),
  `Unemployment Assistance` = c("Flat Rate", "None", "Value 8", "Value 8", "Value 8","Limited"),
  `Univeral Basic Income` = c("Flat Rate", "None", "Value 9", "Value 8", "Value 8","None"),
  `Unemployment Insurance` = c("Relative to prio Income", "Levy", "Value 9", "Value 8", "Value 8","Between X and X")
)

# Generate the table
kable(data, caption = "Example Table Using R")
```
