---
title: "Assistance, Poverty, and the Iron Triangle"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo= FALSE) 
```

```{r}

library(cli)
library(tidyverse)
library(data.table)
library(theme61)
library(tidyr)
library(readxl)
library(gghighlight)
library(readabs)
library(OECD)
library(jsonlite)
library(httr)
library(Synth)
library(knitr)
```

## The Iron Triangle

The key policy framework used to evaluate the trade-offs inherent in setting benefit payment and abatement rates is the **Iron Triangle of Welfare** (cite Henry Tax Review, McClure Report, Welfare Working Group report, and Blundell Keynes Lecture 2021).

```{r}

# Define the vertices of the Iron Triangle
triangle_data <- data.frame(
  x = c(0, 1, 0.5),
  y = c(0, 0, sqrt(3)/2),
  label = c("Adequate Income", "Limited Fiscal Cost", "Work Incentives")
)

ggplot(triangle_data, aes(x = x, y = y)) +
  geom_polygon(fill = "lightblue", color = "black") +
  geom_text(aes(label = label, vjust = ifelse(label == "Work Incentives", -1, 2.5)), 
            hjust = 0.5, size = 8) +
  theme_void() +
  coord_fixed() +
  xlim(-0.5, 1.5) + 
  ylim(-0.5, 1.2)    

```

In terms of objectives, the Iron Triangle is predicated on providing a minimum income standard without "discouraging self-sufficiency" - an objective that is often termed limited disincentives to work.

[@DSS2015]

### Limited Fiscal Cost

\[Fiscal cost\]

\[Also include net social spending - and the point that this is much higher in Australia. If supplementary mechanisms are more necessary, then the idea that the current system is "cost effective" may be overstated\]

### Adequacy

\[Poverty rates\]


```{r}
# https://www.oecd.org/en/data/datasets/income-and-wealth-distribution-database.html

data_url <- "https://sdmx.oecd.org/public/rest/data/OECD.WISE.INE,DSD_WISE_IDD@DF_IDD,1.0/USA+NZL+AUS.A.PR_INC_DISP+PR_INC_MRKT+INC_GROSS_GINI+INC_MRKT_GINI+INC_DISP_GINI..._T.METH2012.D_CUR.?startPeriod=2014&endPeriod=2020"

response <- GET(data_url)

if (response$status_code == 200) {
  # Parse the JSON content
  raw_content <- content(response, "text")
  data <- fromJSON(raw_content, flatten = TRUE)
} else {
  stop("Failed to fetch data from the OECD API")
}

data_sets <- data$dataSets
result <- data.frame()

for (var_name in names(data_sets)) {
  if (grepl("series", var_name) && grepl("observations", var_name)) {
    series_index <- sub("series\\.(\\d+):.*", "\\1", var_name)
    variable_index <- sub("series\\.\\d+:\\d+:(\\d+):.*", "\\1", var_name) # The third number
    sub_variable_index <- sub("series\\.\\d+:\\d+:\\d+:\\d+:\\d+:\\d+:\\d+:\\d+:(\\d+).*", "\\1", var_name) # Ninth number
    obs_number <- sub(".*\\.observations\\.(\\d+)", "\\1", var_name)
    series_id <- sub("series\\.\\d+:(.*)\\.observations\\.\\d+", "\\1", var_name)
    observations <- data_sets[[var_name]]
    value <- observations[[1]][1]
    time <- as.numeric(obs_number)
    result <- rbind(result, data.frame(Series = as.numeric(series_index), Measure = as.numeric(variable_index), Sub_Measure = as.numeric(sub_variable_index) , Time = time, Value = value, SeriesID = series_id))
  }
}

setDT(result)

#result[order(Series,Measure,Time)]
#data$structure[[3]]$series$values

Country_ind <- data.table(Country = data$structure[[3]]$series$values[[1]][,2],Series = c(0,1,2))

result <- Country_ind[result,on=.(Series)]

Indicator_ind <- data.table(Indicator = data$structure[[3]]$series$values[[3]][,2],Measure = seq(0,length(data$structure[[3]]$series$values[[3]][,2])-1))

result <- Indicator_ind[result,on=.(Measure)]

result[, Indicator := fifelse(Sub_Measure == 1, paste0(Indicator, " (50% DI threshold)"),
                      ifelse(Sub_Measure == 2, paste0(Indicator, " (60% DI threshold)"), 
                             Indicator))]

Year_ind <- data.table(Year = seq(2014,2020),Time = seq(0,max(result$Time)))

result <- Year_ind[result,on=.(Time)]

ggplot(result[Sub_Measure > 0],aes(x=Year,y=Value,colour=Country)) + 
  geom_line() +
  theme_e61(legend = "bottom") + 
  scale_y_continuous_e61(limits=c(10,40,5)) +
  labs_e61(title = "Poverty measures",y="") +
  facet_wrap(~Indicator,ncol=2) +
  theme(strip.text = element_text(size = 14))

ggplot(result[Measure %in% c(0,3,4)],aes(x=Year,y=Value,colour=Country)) + 
  geom_line() +
  theme_e61(legend = "bottom") + 
  scale_y_continuous_e61(limits=c(0.2,0.6,0.2)) +
  labs_e61(title = "Inequality",y="") +
  facet_wrap(~Indicator, ncol= 2) +
  theme(strip.text = element_text(size = 14))

```

\- reduction in poverty rate from market to disposable income

\- reduction in Gini from market to gross income

According to [@Whiteford2017] the ratio of social expenditure relative to the reduction in the Gini coefficient can be seen as a measure of **targeting efficiency**. This is the idea of how well a benefit system reduces poverty for a given fiscal outlay.

XX

This relative targeting efficiency is not driven by relative welfare uptake among the unemployed. As highlighted in [@Clarkeetal2023a] XXX

\[Venn diagram and International comparison\]

As a result, the key reason why the Australian system is able to reduce poverty and inequality at a lower cost is due to lower income replacement rates on higher income earners, and relatively elevated support of those out of work for a long period of time.

\[Add something on replacement rates - minimum wage vs average wage\]

\[Add something on replacement rates - short-term vs long-term\]

### Work disincentives

XXX
