---
title: "June 2024 Labour Market Update"
author: "Matt Nolan"
date: "2024-06-13"
output:
  beamer_presentation:
    incremental: true
  ioslides_presentation:
    incremental: true
---

<!-- Initial date: 12/06/2024 -->
<!--  Author: Matt Nolan -->
<!--  Last edit: 12/06/2024 -->
<!--  Purpose: Generate a presentation that goes through important labour market indicators -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

### Read libraries ----
library(readabs)
library(data.table)
library(tidyverse) 
library(ggplot2)
library(knitr)
library(readrba)
library(Hmisc)
library(mFilter)
library(viridis)
library(gridExtra)
library(seasonal)
#library(tinytex) # used when constructing a PDF version, specifically when using beamer for academic presentations

### Set global values ----
last_month <- as.Date("2024-05-01") # Sets a "latest date" used for graphs.

### Construct data of interest
## Base labour market aggregates (quarterly number unemployed, employed, and the unemployment rate) ----
LS <- read_abs(cat_no = "6202.0") 

LS_mini <- LS %>% filter(series_type == "Seasonally Adjusted") %>% filter(table_title == "Table 1. Labour force status by Sex, Australia - Trend, Seasonally adjusted and Original") %>% filter(series == "Employed total ;  Persons ;" | series == "Unemployed total ;  Persons ;"  | series == "Unemployment rate ;  Persons ;") %>% select("date","series","value")

LS2 <- pivot_wider(LS_mini,id_cols = "date",names_from = "series",values_from = "value")
colnames(LS2) <- c("date","Employed","Unemployed","UR_ABS")

LS2 <- LS2 %>% mutate(UR_Derived = Unemployed/(Employed + Unemployed),QUnemp = frollmean(Unemployed,n=3,fill=NA,align="right"),QEmp =frollmean(Employed,n=3,fill=NA,align="right"),UR = UR_ABS) 

LSQ <- LS2[seq(5,nrow(LS2),3),] %>% select(date,QUnemp,QEmp,UR)


## Labour market tightness: Use the ABS data - this is quarterly ----

vac_full <- read_abs(cat_no = "6354.0")

vac <- vac_full %>% filter(series == "Job Vacancies ;  Australia ;") %>% filter(grepl('TABLE 1',table_title)) %>% filter(series_type == "Seasonally Adjusted")

vac$date <- vac$date %m+% months(1) -14

vac <- vac %>% select(date,value)
colnames(vac) <- c("date","vacancy")

# Need to impute the data that was not collected for August 2008 to August 2009 based on a regression of the log difference in vacancies against the change in the unemployment rate from pre-May 2008. 

vacdf <- left_join(vac,LSQ,by="date") %>% select(date,vacancy,QUnemp) %>% mutate(time = seq(1,nrow(vac),by=1))

vac_imp <- aregImpute(~vacancy + QUnemp + time,data=vacdf %>% filter(date <= "2020-01-01"))

vacdf[is.na(vacdf$vacancy),]$vacancy <- vac_imp$imputed$vacancy[nrow(vac_imp$imputed$vacancy),]

plot(vacdf$vacancy)
plot(vacdf$QUnemp)

vu <- left_join(vacdf,LSQ, by="date")

v <- vacdf %>% transmute(date=date,v=vacancy,Unemp=QUnemp)
setDT(v)
v <- melt(v,id = "date")

vu <- vu %>% transmute(date=date,vu = vacancy/(QUnemp.x*1000))
setDT(vu)

average_vu_pc <- mean(vu[date > as.Date("1999-12-01") & date < as.Date("2020-01-01")]$vu, na.rm = TRUE) * 1000 # Average tightness ratio between 2000 and 2019. Look at pre-COVID, given the COVID specific patterns.

## Add industry vacancy data - only use from 2013 onwards ----

vac_ind <- vac_full %>% filter(startsWith(table_title, "TABLE 4") & !startsWith(series, "Standard Error")) %>% mutate(industry = str_extract(series, ";\\s*(.*?)\\s*;")) %>% mutate(industry = str_trim(str_remove_all(industry, ";"))) %>% select(date,value,industry)
setDT(vac_ind)
setkey(vac_ind,industry,date)

vac_ind[,annual_vac := frollmean(value,12,fill=NA,align="right")]

#ggplot(vac_ind[industry %in% unique(vac_ind$industry)[c(9,17)]],aes(x=date,y=annual_vac,colour=industry)) + geom_line() # Illustrates an issue with the pre 2013 data (18 even more so).
# temp <- vac_ind[industry != "Total All Industries",.(value = sum(value),industry = "total_implied",annual_vac = sum(annual_vac,na.rm = TRUE)),by=date]
# vac_ind <- rbind(vac_ind,temp)
# ggplot(vac_ind[date > as.Date("2012-12-30")],aes(x=date,y=annual_vac,colour=industry)) + geom_line()

vac_ind <- vac_ind[date > as.Date("2012-12-30")]

vac_ind[,agg_ind := fcase(industry %in% c("Mining","Construction","Manufacturing","Electricity, Gas, Water and Waste Services","Transport, Postal and Warehousing","Wholesale Trade"),"Primary and Secondary",
                          industry %in% c("Retail Trade","Accommodation and Food Services","Rental, Hiring and Real Estate Services" ),"Sales",
                          industry %in% c("Administrative and Support Services","Financial and Insurance Services" ,"Professional, Scientific and Technical Services","Information Media and Telecommunications"),"Private Services",
                          industry %in% c("Education and Training","Health Care and Social Assistance","Public Administration and Safety"),"Public Services",
                          industry %in% c("Arts and Recreation Services","Other Services"),"Other Services",
                          industry %in% c("Total All Industries","total_implied"),"Total",
                          default = NA)]

vac_ind[,agg_total := sum(value),by=.(agg_ind,date)]
vac_ind[,agg_ann_total := sum(annual_vac),by=.(agg_ind,date)]

ggplot(vac_ind[date > as.Date("2012-12-30") & !industry %in% c("Total All Industries","total_implied"  )],aes(x=date,y=annual_vac,colour=industry)) + geom_line()

ggplot(vac_ind[date > as.Date("2012-12-30") & !agg_ind %in% c("Total")],aes(x=date,y=agg_ann_total,colour=agg_ind)) + geom_line()
ggplot(vac_ind[date > as.Date("2012-12-30") & !agg_ind %in% c("Total")],aes(x=date,y=agg_total,colour=agg_ind)) + geom_line()

# Make industry vacancy rates

LS_ind <- read_abs(cat_no = "6291.0.55.001") %>% filter(table_title == "Table 04. Employed persons by Industry division of main job (ANZSIC) - Trend, Seasonally adjusted, and Original") %>% filter(series_type == "Original")
setDT(LS_ind)

LS_ind <- LS_ind[series != "Employed total ;"][, industry := str_remove(series, " ;  Employed total ;")][industry != "Agriculture, Forestry and Fishing"]

LS_ind[,agg_ind := fcase(industry %in% c("Mining","Construction","Manufacturing","Electricity, Gas, Water and Waste Services","Transport, Postal and Warehousing","Wholesale Trade"),"Primary and Secondary",
                         industry %in% c("Retail Trade","Accommodation and Food Services","Rental, Hiring and Real Estate Services" ),"Sales",
                         industry %in% c("Administrative and Support Services","Financial and Insurance Services" ,"Professional, Scientific and Technical Services","Information Media and Telecommunications"),"Private Services",
                         industry %in% c("Education and Training","Health Care and Social Assistance","Public Administration and Safety"),"Public Services",
                         industry %in% c("Arts and Recreation Services","Other Services"),"Other Services",
                         industry %in% c("Total All Industries","total_implied"),"Total",
                         default = NA)]

LS_ind <- LS_ind[date > as.Date("2012-12-30"),.(date,value,industry,agg_ind)][,date := date + 14][,agg_emp := sum(value),by=.(date,agg_ind)]

ind_vac_emp <- vac_ind[LS_ind,on=.(date,industry)]

# Note. This overstates the vacancy rate, as it doesn't include the jobs of multiple job holders. Instead of employment we would want to use filled jobs.
vac_rate_ind <- ind_vac_emp[,.(vac_rate = agg_total/(agg_emp+agg_total)),by=.(date,agg_ind)]

### Construct alternative vacancy rate with labour accounts measures (filled jobs, total jobs, etc) [preferred measure for industry level] ----

labour_acc_full <- read_abs(cat_no = "6150.0.55.003")
setDT(labour_acc_full)
unique(labour_acc_full$table_title)

labour_acc_full[, industry := str_match(table_title, "Table \\d+\\. (.*?) - Trend, Seasonally Adjusted and Original")[, 2]]
labour_acc_full <- labour_acc_full[!is.na(industry)]

# These accounts include values for Agriculture
labour_acc_full[,agg_ind := fcase(industry %in% c("Agriculture, Forestry and Fishing (A)","Mining (B)","Construction (E)","Manufacturing (C)","Electricity, Gas, Water and Waste Services (D)","Transport, Postal and Warehousing (I)","Wholesale Trade (F)"),"Primary and Secondary",
                                  industry %in% c("Retail Trade (G)","Accommodation and Food Services (H)","Rental, Hiring and Real Estate Services (L)"),"Sales",
                                  industry %in% c("Administrative and Support Services (N)","Financial and Insurance Services (K)" ,"Professional, Scientific and Technical Services (M)","Information Media and Telecommunications (J)"),"Private Services",
                                  industry %in% c("Education and Training (P)","Health Care and Social Assistance (Q)","Public Administration and Safety (O)"),"Public Services",
                                  industry %in% c("Arts and Recreation Services (R)","Other Services (S)"),"Other Services",
                                  industry %in% c("Total All Industries"),"Total",
                                  default = NA)]

labour_acc <- labour_acc_full[!is.na(agg_ind)]

unique(labour_acc[table_title == "Table 1. Total All Industries - Trend, Seasonally Adjusted and Original" & series_type=="Original" & date > as.Date("2012-12-30") & startsWith(series,"Jobs")]$series)

ggplot(labour_acc[table_title == "Table 1. Total All Industries - Trend, Seasonally Adjusted and Original" & series_type=="Seasonally Adjusted" & date > as.Date("2012-12-30") & startsWith(series,"Jobs; Proportion of vacant jobs ;")],aes(x=date,y=value,colour = agg_ind)) + geom_line() + theme_classic()

ggplot(labour_acc[table_title == "Table 1. Total All Industries - Trend, Seasonally Adjusted and Original" & series_type=="Seasonally Adjusted" & date > as.Date("1999-12-30") & startsWith(series,"Jobs; Proportion of vacant jobs ;")],aes(x=date,y=value,colour = agg_ind)) + geom_line() + theme_classic()


labour_acc <- labour_acc[(startsWith(series,"Jobs; Total jobs") | startsWith(series,"Jobs; Job vacancies")) & unit == "000" & date > "2012-12-30" & series_type == "Seasonally Adjusted",.(date,industry,agg_ind,value,series)]

labour_acc[, series2 := str_extract(series, "^[^;]+;[^;]+")]

labour_acc <- dcast(labour_acc,date+industry+agg_ind~series2)

labour_acc[,vac_rate := `Jobs; Job vacancies `/`Jobs; Total jobs `][,agg_vac_rate := sum(`Jobs; Job vacancies `)/sum(`Jobs; Total jobs `),by=.(date,agg_ind)]

## Multiple job holding rate. ----

mult_job <- labour_acc_full[series == "Persons; Rate of multiple job holding (proportion of employed persons) ;  Australia ;  Total all industries ;" & series_type == "Seasonally Adjusted"][,.(date,mult_job_rate = value)] 

ggplot(mult_job,aes(x=date,y=mult_job_rate)) + geom_line()

## Participation rates ----

LS_part <- LS %>% filter(series_type == "Seasonally Adjusted") %>% filter(table_title == "Table 1. Labour force status by Sex, Australia - Trend, Seasonally adjusted and Original")  %>% filter(series == "Participation rate ;  Persons ;") %>% select("date","series","value")
setDT(LS_part)

avg_PR <- mean(LS_part[date > as.Date("2004-12-30") & date < as.Date("2020-01-01")]$value)

LS_part[,avg_PR := avg_PR][,dPR := value - avg_PR]

ggplot(LS_part[date > as.Date("2012-12-30")],aes(x=date,y=value)) + geom_line() + geom_hline(yintercept = avg_PR,colour="red",linetype="dashed")


### Earnings ----

# By industry (COE and COE per job) - decided to exclude for time.

# By WPI
wpi <- read_abs("6345.0")
setDT(wpi)

wpi_SA <- wpi[series_type == "Seasonally Adjusted" & data_type == "INDEX"][, series2 := sapply(str_split(series, ";\\s+"), function(x) x[4])]

ggplot(wpi_SA[],aes(x=date,y=value,colour=series2)) + geom_line()

unique(wpi[str_detect(series, regex("year", ignore_case = TRUE)) & 
             str_detect(series, regex("Total", ignore_case = TRUE)) & 
             str_detect(series, regex("All industries", ignore_case = TRUE)) &
             str_detect(series, regex("Australia", ignore_case = TRUE)) &
             str_detect(series, regex("Private and Public", ignore_case = TRUE)) &
             series_type == "Original" & 
             data_type == "PERCENT" & 
             date > as.Date("2012-12-30")]$series)

ann_wpi_growth <- wpi[series %in% c("Percentage Change from Corresponding Quarter of Previous Year ;  Total hourly rates of pay excluding bonuses ;  Australia ;  Private and Public ;  All industries ;","Percentage Change from Corresponding Quarter of Previous Year ;  Total hourly rates of pay excluding bonuses ;  Australia ;  Public ;  All industries ;","Percentage Change from Corresponding Quarter of Previous Year ;  Total hourly rates of pay excluding bonuses ;  Australia ;  Private ;  All industries ;") & series_type == "Original" & 
      data_type == "PERCENT" & 
      date > as.Date("2012-12-30") & table_no == "634501"]

ann_wpi_growth <- ann_wpi_growth[, series2 := sapply(str_split(series, ";\\s+"), function(x) x[4])][,.(date,value,series2)]

cpi <- read_abs(cat_no = "6401.0")
setDT(cpi)

cpi_growth <- cpi[table_title == "TABLES 1 and 2. CPI: All Groups, Index Numbers and Percentage Changes" & unit == "Percent" & series == "Percentage Change from Corresponding Quarter of Previous Year ;  All groups CPI ;  Australia ;" & date > as.Date("2012-12-30")][,.(date,cpi_value = value)]

cpi_growth <- cpi_growth[ann_wpi_growth[series2 == "Private and Public "],on=.(date)][,RW := value - cpi_value]

ggplot(cpi_growth,aes(x=date,y=RW)) + geom_line() + geom_hline(yintercept = 0)

# Add average RW growth 2004 - 2019. For this get December 2019 WPI and CPI, and December 2004 WPI and CPI, and calculate annualised growth.

cpi_for_avg <- cpi[(date == "2004-12-01" | date == "2019-12-01") & table_title == "TABLES 1 and 2. CPI: All Groups, Index Numbers and Percentage Changes" & series == "Index Numbers ;  All groups CPI ;  Australia ;"][,.(date,cpi_value = value)]
setDT(cpi_for_avg)

wpi_for_avg <- wpi[series == "Quarterly Index ;  Total hourly rates of pay excluding bonuses ;  Australia ;  Private and Public ;  All industries ;" & (date == "2004-12-01" | date == "2019-12-01") & series_type == "Original" & table_title == "Table 1. Total Hourly Rates of Pay Excluding Bonuses: Sector, Original, Seasonally Adjusted and Trend"][,.(date,wpi_value = value)]
setDT(wpi_for_avg)

avg_RW <- cpi_for_avg[wpi_for_avg,on=.(date)][,RW := wpi_value/cpi_value]

avg_RW_growth <- (avg_RW$RW[2]/avg_RW$RW[1])^(1/15)-1

cpi_growth[,dRW := RW - avg_RW_growth*100]

ggplot(cpi_growth,aes(x=date,y=dRW)) + geom_line() + geom_hline(yintercept = 0)

### Flow data ----
## EU, UE, J2Js

dfgf <- read_lfs_grossflows(weights = "previous")
setDT(dfgf)
dfgf[,year := year(date)][,month := month(date)]

# March quarters
#gf <- dfgf[month < 4] %>% group_by(year,lfs_current,lfs_previous) %>% summarise(number = sum(persons))

#ggplot(gf %>% filter(year > "1999" & lfs_current == "Unemployed" & (lfs_previous == "Employed full-time" | lfs_previous == "Employed part-time")), aes(x=year,y=number,colour=lfs_previous)) + geom_line() # EU

#ggplot(gf %>% filter(year > "1999" & lfs_previous == "Unemployed" & (lfs_current == "Employed full-time" | lfs_current == "Employed part-time")), aes(x=year,y=number,colour=lfs_current)) + geom_line() #UE

# Add a quarter indicator and seasonally adjust to make quarterly series
dfgf <- dfgf %>%
  mutate(quarter = ceiling(month / 3))

gf <- dfgf %>%
  group_by(year, quarter, lfs_current, lfs_previous) %>%
  summarise(number = sum(persons), .groups = 'drop') %>% filter(lfs_current == "Unemployed" & (lfs_previous == "Employed full-time" | lfs_previous == "Employed part-time"))

gf <- gf %>%
  mutate(date = as.Date(paste(year, (quarter - 1) * 3 + 1, 1, sep = "-")))

gf <- gf %>%
  group_by(lfs_current, lfs_previous) %>%
  mutate(number_sa = {
    ts_data <- ts(number, start = c(min(year), min(quarter)), frequency = 4)
    final(seas(ts_data))
  }) %>%
  ungroup()
setDT(gf)

ggplot(gf[date < as.Date("2024-04-01")],aes(x=date,y=number_sa,colour=lfs_previous)) + geom_line()

# Annual mobility indicators - only up to February 2023 so did not include
#PJSM <- read_abs(cat_no = "6223.0")


### Productivity ----

GDP <- read_abs(cat_no = "5206.0") %>% filter(series_type == "Seasonally Adjusted") %>% filter(series == "Gross domestic product: Chain volume measures ;") %>% filter(table_title == "Table 1. Key National Accounts Aggregates") %>% select("date","series","value")

Hours <- read_abs(cat_no = "5206.0") %>% filter(series_type == "Seasonally Adjusted") %>% filter(series == "Hours worked: Index ;") %>% filter(grepl('Table 1',table_title)) %>% drop_na %>% filter(unit == "Index Numbers") %>% select("date","value")

colnames(Hours) <- c("date","Hours")

LP <- left_join(LSQ,GDP,by="date") %>% mutate(Labour_Prod = value/QEmp) %>% select(date,Labour_Prod) %>% drop_na
LP$Labour_Prod <- LP$Labour_Prod/LP[LP$date == "1980-03-01",]$Labour_Prod

LP_hour <- left_join(Hours,GDP,by="date")%>% mutate(Hour_Labour_Prod = value/Hours) %>% select(date,Hour_Labour_Prod) %>% drop_na
LP_hour$Hour_Labour_Prod <- LP_hour$Hour_Labour_Prod/LP_hour[LP_hour$date == "1980-03-01",]$Hour_Labour_Prod

hp_lp_result <- hpfilter(LP$Labour_Prod, freq = 1600)

LP$Trend <- hp_lp_result$trend

hp_lph_result <- hpfilter(LP_hour$Hour_Labour_Prod, freq = 1600)

LP_hour$Hour_Trend <- hp_lph_result$trend

LP_full <- left_join(LP,LP_hour,by="date")

LP_full <- LP_full %>% pivot_longer(!date)

ggplot(LP_full,aes(x=date,y=value[,1],colour=name)) + geom_line() #+ plot_label("Hourly Productivity",y=1.8,x="1985-03-01",colour=palette_e61(4)[2]) + plot_label("Output per employee",y=1.6,x="1985-03-01",colour=palette_e61(4)[3]) + labs_e61(title = "Labour Productivity", subtitle = "Index, 1980 = 1, HP smoothing parameter = 1600",y="",x="") + scale_y_continuous_e61(limits = c(0.9,2,0.2),y_top = FALSE)

# LP growth

setDT(LP)
LP$Trend <- data.table(LP$Trend)
LP[,lag_Trend := shift(Trend,4)][,LP_trend_growth := Trend/lag_Trend - 1]
LP[,lag_LP := shift(Labour_Prod,4)][,LP_growth := Labour_Prod/lag_LP - 1]

ggplot(LP[date > as.Date("2012-12-30")],aes(x=date,y=LP_trend_growth)) + geom_line()
ggplot(LP[date > as.Date("2012-12-30")],aes(x=date,y=LP_growth)) + geom_line()

ggplot(melt(LP[date > as.Date("2012-12-30"),.(date,LP_growth,LP_trend_growth)],id="date"),aes(x=date,y=value,colour=variable)) + geom_line()

### Unemployment rate ----
# 
# UR <- LSQ %>% select(date,UR)
# setDT(UR)

UR <- LS %>% filter(series_type == "Seasonally Adjusted" | series_type == "Trend") %>% filter(table_title == "Table 1. Labour force status by Sex, Australia - Trend, Seasonally adjusted and Original")  %>% filter(series == "Unemployment rate ;  Persons ;") %>% select("date","series_type","value")
setDT(UR)

avg_UR <- mean(UR[date > as.Date("2004-12-30") & date < as.Date("2020-01-01")]$value)

```

# Current Australian Labour Market

## Outline

Goal of this presentation is to understand the current state of the (post-COVID) Australian labour market.  Key takeaways:

- Labour market remains tight - but signs of weakness are appearing.
- However, there are significant wage and productivity puzzles.

These two facts suggest that labour market search flows are likely to cool - but the outlook for wages remain uncertain.


## Labour market is softening - but remains tight

```{r URPR, echo = FALSE}

UR_plot <- ggplot(UR[date > as.Date("2012-12-30")], aes(x = date, y = value, colour = series_type)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = avg_UR, linetype = "dashed", color = "black") +
  labs(
    title = "Unemployment rate",
    subtitle = "Monthly, seasonally adjusted",
    y="",
    x="",
    caption = "ABS"
  ) +
  theme_classic() +
  scale_colour_manual(values = c("Seasonally Adjusted"="red","Trend"="blue")) +
  theme(legend.position = "none") +
  annotate("text", x = as.Date("2013-01-01"), y = 7.2, label = "Seasonally Adjusted", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 7.2 - 0.5, label = "Trend", color = "blue", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = avg_UR - 0.5, label = "2005-2019 UR Average", color = "black", hjust = 0) + scale_y_continuous(labels = scales::percent_format(scale = 1),limits = c(3,8))


PR_plot <- ggplot(LS_part[date > as.Date("2012-12-30")],aes(x=date,y=value,colour=series)) + geom_line(linewidth = 1) + geom_hline(yintercept = avg_PR,linetype = "dashed") + 
  theme_classic() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("Participation rate ;  Persons ;" = "red")) +
  annotate("text", x = as.Date("2013-01-01"), y = avg_PR + 1.2, label = "2005-2019 PR Average", color = "black", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = avg_PR + 1.7, label = "Participation Rate", color = "red", hjust = 0) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),limits = c(62,68)) +
  labs(title = "Participation Rate",subtitle = "Monthly, seasonally adjusted",y="",x="",caption="ABS")

#grid.arrange(UR_plot, PR_plot, tight_plot, tight_plot, ncol = 2, nrow = 2) # Will add participation, J2J, UR
grid.arrange(UR_plot, PR_plot, ncol = 2, nrow = 1) 


```

## Labour market is softening - but remains tight

```{r tightness, echo = FALSE}

flow_plot <- ggplot(gf[date < as.Date("2024-04-01") & date > as.Date("2012-12-30")],aes(x=date,y=number_sa,colour=lfs_previous)) + geom_line(linewidth = 1) +
  theme_classic() +
  scale_colour_manual(values = c("Employed full-time" = "red","Employed part-time" = "blue")) +
  theme(legend.position = "none") + labs(title = "Transition to Unemployment",subtitle="Quarterly (000s), seasonally adjusted",y="",x="",caption="ABS") +
  annotate("text", x = as.Date("2013-01-01"), y = 180, label = "From full-time", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 210, label = "From part-time", color = "blue", hjust = 0)
  

tight_plot <- ggplot(vu[date > as.Date("1999-12-01")],aes(x=date,y=vu*1000)) + geom_line(linewidth = 1,colour="red") + theme_classic() + labs(
  title = "'Tightness' highly elevated",
  subtitle = "Vacancies per unemployed person",
  y="",
  x="",
  caption = c("ABS")
) + geom_hline(yintercept = average_vu_pc, linetype = "dashed", color = "black") +
  theme(legend.position = "none") +
  annotate("text", x = as.Date("2000-01-01"), y = 0.6, label = "V/U ratio", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2000-01-01"), y = 0.5, label = "2000-2019 average", color = "black", hjust = 0)
  

grid.arrange(flow_plot, tight_plot, ncol = 2, nrow = 1) 


```

## Firm demand the main driver

```{r vacancy-unemployment,echo=FALSE}
vu_plot <- ggplot(v[date > as.Date("1999-12-01")],aes(x=date,y=value,colour=variable)) + geom_line(linewidth=1) + theme_classic() + labs(
  title = "Vacancies and the unemployed",
  subtitle = "Vacancy and unemployment numbers (000s)",
  y="",x="",
  caption = c("ABS"),
  footnote = "Job Vacancies Survey"
) +  scale_colour_manual(values = c("v" = "red","Unemp" = "blue")) + theme(legend.position = "none") +
  annotate("text", x = as.Date("2000-01-01"), y = 300, label = "Vacancies", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2000-01-01"), y = 800, label = "Unemployed", color = "blue", hjust = 0) 
  

v_ind_plot <- ggplot(labour_acc,aes(x=date,y=agg_vac_rate,colour=agg_ind,linetype=agg_ind,linewidth = agg_ind)) + geom_line()+ theme_classic() + labs(
  title = "Vacancy Rate by Industry",
  subtitle = "Vacancies divided by desired jobs",
  y="",
  caption = c("ABS"),
  footnote = "Based on experimental Labour Accounts. Similar trends hold using the general vacancy and employment data."
) + scale_y_continuous(labels = scales::percent_format()) + #scale_colour_viridis(discrete = TRUE,option="D")
  scale_colour_manual(values = c("Total"="black","Sales" = "green","Private Services"="blue","Primary and Secondary" = "cyan1","Other Services" = "pink","Public Services"="red")) +
  scale_linetype_manual(values = c("Total"=1,"Sales" = 1,"Private Services"=2,"Primary and Secondary" = 2,"Other Services" = 2,"Public Services"=1)) +
  scale_linewidth_manual(values = c("Total"=1,"Sales" = 0.2,"Private Services"=0.2,"Primary and Secondary" = 0.2,"Other Services" = 0.2,"Public Services"=0.2)) + theme(legend.position = "none") +
  annotate("text", x = as.Date("2013-01-01"), y = 0.04, label = "Private Services", color = "blue", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.038, label = "Sales", color = "green", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.036, label = "Other Services", color = "pink", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.034, label = "Primary and Secondary", color = "cyan1", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.032, label = "Public Services", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.030, label = "Total", color = "black", hjust = 0) 
  

grid.arrange(vu_plot, v_ind_plot, ncol = 2, nrow = 1)

```

## But why have firms been listing vacancies with weak productivity?

```{r J2J,echo=FALSE}

ggplot(melt(labour_acc[industry == "Total All Industries" & date > as.Date("2012-12-30")][,.(date,vac_rate)][LP[date > as.Date("2012-12-30"),.(date,LP_growth,LP_trend_growth)],on=.(date)],id="date"),aes(x=date,y=value,colour=variable,linetype=variable)) + geom_line() + theme_classic() +theme(legend.position = "none") +
  labs(title = "Labour Productivity and Vacancy Rates",subtitle="Quarterly value",x="",y="",caption="ABS") +
  scale_colour_manual(values=c("vac_rate"="red","LP_trend_growth"="blue","LP_growth"="aquamarine2")) +
  scale_linetype_manual(values =c("vac_rate"=1,"LP_trend_growth"=1,"LP_growth"=2)) +
  annotate("text", x = as.Date("2013-01-01"), y = -0.005, label = "Vacancy Rate", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = -0.015, label = "Labour Productivity Growth", color = "aquamarine2", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = 0.035, label = "Trend Labour Productivity Growth", color = "blue", hjust = 0) + scale_y_continuous(labels=scales::percent_format()) + geom_hline(yintercept = 0)



```


## And why are households so willing to work with lower real wages?

```{r puzzles,echo=FALSE}

graph8_dt <- LS_part[date > as.Date("2012-12-30"),.(date,dPR)][cpi_growth,on=.(date)][,.(date,dPR,dRW)]

ggplot(melt(graph8_dt,id = "date"),aes(x=date,y=value,colour=variable))+geom_line(linewidth=1) + geom_hline(yintercept = 0) + theme_classic() + labs(title="Worker outcomes relative to pre-COVID average",subtitle="Difference from 2005-2019 average",y="",x="",caption="ABS") +theme(legend.position = "none") +
  scale_colour_manual(values=c("dPR"="red","dRW"="blue")) + 
  annotate("text", x = as.Date("2013-01-01"), y = 1, label = "Participation Rate", color = "red", hjust = 0) +
  annotate("text", x = as.Date("2013-01-01"), y = -2, label = "Real Wage Growth", color = "blue", hjust = 0) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),limits = c(-6,2))
  


#ggplot(mult_job,aes(x=date,y=mult_job_rate)) + geom_line()


```


# Explaining these puzzles


## Potential explanations 

Three sets of explanations. All suggest that labour market turnover is set to cool. However, have opposite implications for measured wages.

- **Reallocation**: Change in job requirements post-COVID - wages will rise as job tenure rises.
- **Amenities**: Households willing to trade wages for flexibility - no pressure on wages.
- **Cost of living**: Rising prices encouraging people to work - "money illusion" will pass leading to wage gains.


# Summing up

## Summary

There are signs the labour market is cooling, with search behaviour declining by firms while a rising number of individuals are finding themselves out of work.

But in historic terms the labour market is still very tight.

Labour market turnover will continue to cool, but the outlook for wages remains uncertain.
